---
title: "myBookQ"
author: "Alicia Brown"
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    orientation: rows
    social: [ "twitter", "facebook", "linkedin"]
    source_code: "https://github.com/aliciatb/mybookq"
    favicon: favicon-q.png
runtime: shiny
---

```{r setup}
#options(shiny.sanitize.errors = T)

library(dplyr)
library(ggplot2)
library(lubridate)
library(flexdashboard)
library(httr)
library(jsonlite)
library(reshape2)
library(scales)
library(stringr)
library(tidyr)
library(timevis)
library(tidytext)
library(wordcloud)
require(wesanderson)

favorite_authors <- "Louise Penny,Fred Vargas,Kate Atkinson,Nancy Atherton,Jussi Adler-Olsen,Val McDermid,Tana French"
other_authors <- "Val McDermid,Lisa Scottoline,Jacqueline Winspear,Stella Rimington,Susan Elia MacNea,Denisa Mina,Val McDermid,Tana French,Deborah Crombie,Mike Lawson,Arnaldur Indridason,Susan Hill,Elly Griffiths,M. L. Longworth,Helene Tursten,Camilla Lackberg,David Lagercrantz,Nevada Barr"
authors <- paste0(favorite_authors)

positive_continuing_series_words <- "return,returns,returning,next,series,continue,continuing,gamache,brodie,adamsberg,evangelists,finch"
negative_continuing_series_words <- "stand-alone"
positive_words <- "cozy,dog,cat,village,eccentric,kind,thoughtful,sympathy"
negative_words <- "psychological,sociopath,psychopath,serial,hunting,cruel,mean"
```

```{r authorURL}
authorURL <- function(author){
  google_key <- 'AIzaSyAwJfF6E3x_UkuAj4ahufzJKwOHZIaEql0'
  
  # currently pulling last author, need to iterate and rbind
  author_name <- str_split(author," ", simplify = TRUE)
  
  author_first <- author_name[1]
  author_last <- author_name[2]
  
  order_by <- 'newest'
  # https://www.googleapis.com/books/v1/volumes?q=inauthor:Fred%20Vargas&langRestrict=en&orderBy=newest&printType=books&maxresults=40&prettyPrint=true&key=AIzaSyAwJfF6E3x_UkuAj4ahufzJKwOHZIaEql0
  url <- paste0("https://www.googleapis.com/books/v1/volumes?q=inauthor:",author_first,"+",author_last,"&langRestrict=en&orderBy=",order_by,"&printType=books&maxresults=40&prettyPrint=true&key=",google_key)
  url
}
```

```{r anticipationScores}
anticipationScores <- function(){
  # for debug, commment out inputs
  positive_words <- input$description_positive
  negative_words <- input$description_negative
  positive_continuing_series_words <- input$description_positive_continuing_series
  negative_continuing_series_words <- input$description_negative_continuing_series
  
  # combine positive and continuing
  positive_words <- paste0(positive_words,",",positive_continuing_series_words)
  negative_words <- paste0(negative_words,",",negative_continuing_series_words)
  
  positive_scores <- tibble(word = positive_words)
  negative_scores <- tibble(word = negative_words)

  positive_scores <- positive_scores %>%
    mutate(word = strsplit(word, ",")) %>%
    unnest(word) %>%
    mutate(score = 1, sentiment = "positive")
  
  negative_scores <- negative_scores %>%
    mutate(word = strsplit(word, ",")) %>%
    unnest(word) %>%
    mutate(score = -1, sentiment = "negative")
    
  scores <- bind_rows(positive_scores, negative_scores)
  scores
}

```

```{r}
# Reactive data available for all shiny modules
getData <- reactive({
  # 1. Get the comma delimited string of authors from input$author_query
  
  date_from <- input$dateRange[1]
  date_to = input$dateRange[2]
  author_query <- c(input$author_query)
  
  # for debug
  # author_query <- favorite_authors
  # date_from <- Sys.Date() - 60
  # date_to <- Sys.Date() + 365
  
  authors <- str_split(author_query,",",simplify = T)
  
  books <- tibble()
  
  # 2. Get books for each author, filter by author name, combine data
  for(author in authors){
    # debug
    # author <- 'Louise Penny'
    author_url <- authorURL(author)
    data <- fromJSON(URLencode(author_url),simplifyDataFrame = TRUE)
    
    volume_info <- data$items$volumeInfo
    access_info <- data$items$accessInfo
    search_info <- data$items$searchInfo
    sale_info <- data$items$saleInfo
    
    # debug
    # test <- raw_data %>%
    #   filter(authors == 'Louise Penny')
    
    if (is.null(volume_info) == FALSE){
      raw_data <- cbind(volume_info,sale_info,search_info) %>%
        filter(authors == author) %>%
        # adding author here since results returns as list
        mutate(name = author) %>%
        # handle year and year-month dates
        mutate(
          publishedDate = case_when(
          str_length(publishedDate) == 4 ~ paste0(publishedDate,"-01-01"),
          str_length(publishedDate) == 7 ~ paste0(publishedDate,"-01"),
          TRUE ~ publishedDate)) %>%
        # filter by date control
        filter(publishedDate >= date_from) %>%
        filter(publishedDate <= date_to) %>%
        select(title,name,publishedDate,description,saleability) %>%
        mutate(description_length = str_length(description)) %>%
        filter(description_length > 0) %>%
        arrange(desc(publishedDate))
      
      books <- rbind(books, raw_data)
    }
  }
  
  books <- books %>%
    select(title,name,publishedDate,description) %>%
    group_by(title,name) %>%
    summarize(publishedDate = min(publishedDate),
              description = max(description))
  
  books

})
```

Sidebar {.sidebar}
======================================================================

Check for new  mysteRies!

```{r}
textAreaInput("author_query", label = "Favorite Authors:", value = authors, rows=3, resize = "vertical")
```

```{r}
  dateRangeInput('dateRange',
      label = 'Date range:',
      start = Sys.Date() - 30, end = Sys.Date() + 365
    )
```

```{r}
textAreaInput("description_positive", label = "Positive Keywords:", value = positive_words, rows=2, resize = "vertical")
```

```{r}
textAreaInput("description_negative", label = "Negative Keywords:", value = negative_words, rows=2, resize = "vertical")
```

```{r}
textAreaInput("description_positive_continuing_series", label = "Continuing Keywords:", value = positive_continuing_series_words, rows=2, resize = "vertical")
```

```{r}
textAreaInput("description_negative_continuing_series", label = "New Series Keywords:", value = negative_continuing_series_words, rows=2, resize = "vertical")
```

Author: [Alicia Brown](https://msaliciabrown.com)

Books
======================================================================

Row
-----------------------------------------------------------------------

### Available books {.value-box}

```{r available_book_count}
renderValueBox({
  data_view <- getData()
  # for debug
  # data_view <- books
  
  if(length(data_view) > 0){
    data_view <- data_view %>%
      filter(publishedDate <= Sys.Date()) %>%
      distinct(name,title)
      
    book_count <- nrow(data_view)
    
    valueBox(book_count, 
             icon = "fa-book",
             color = "info")
  }
})
```

### Upcoming books {.value-box}

```{r upcoming_book_count}
renderValueBox({
  data_view <- getData()
  # for debug
  # data_view <- books
  
  if(length(data_view) > 0){
    data_view <- data_view %>%
      filter(publishedDate < Sys.Date() + 30) %>%
      filter(publishedDate > Sys.Date()) %>%
      distinct(name,title)
      
    book_count <- nrow(data_view)
    
    valueBox(book_count, 
             icon = "fa-calendar",
             color = "info")
  }
})
```

### Anticipation index {.gauge}

```{r anticipation_index}
renderGauge({
  
  data_view <- getData()  
  anticipation <- anticipationScores()
  
  if(length(data_view) > 0){
    
     book_description_anticipation <- separate_rows(data_view, description, sep = "\\.", convert = FALSE) %>%
        select(title, name, description) %>%
        drop_na(description) %>%
        group_by(title,name) %>%
        mutate(description_line = row_number()) %>%
        ungroup() %>%
        unnest_tokens(word, description) %>%
        inner_join(anticipation) %>%
        summarise(anticipation = sum(score))
     
     books_score <- book_description_anticipation$anticipation
    
    gauge(books_score, min = -10, max = 10, gaugeSectors(
      success = c(5, 15), warning = c(-5, 14), danger = c(-55, -5)
    ))
  }
})
```

Row
----------------------------------------------------------------------------

### Upcoming books (next 30 days) and recently published (last 14 days)

```{r upcoming_time_vis, fig.height=3}
renderTimevis({
  data_view <- getData()
  # for debug
  # data_view <- books
  
  if(length(data_view) > 0){
    data_view <- data_view %>%
      filter(publishedDate < Sys.Date() + 30) %>%
      filter(publishedDate > Sys.Date() - 14) %>%
      distinct(name,title,publishedDate) %>%
      mutate(name_title = paste(title,"-",name))
  }
  
  if(nrow(data_view) > 0)
  {
    data <- data.frame(
         id      = 1:nrow(data_view),
         content = c(data_view$name_title),
         start = c(data_view$publishedDate)
    )
  timevis(data)
  }
})
```


Row
-----------------------------------------------------------------------

### All search results
    
```{r new_books, fig.width=14, fig.height=10}

renderTable({
  
  data_view <- getData()
  # for debug
  # data_view <- books
  
  # sentiments
  afinn <- get_sentiments("afinn")
  # anticipation score
  anticipation <- anticipationScores()
   
  if(length(data_view) > 0){
    
    book_description_sentiments <- separate_rows(data_view, description, sep = "\\.", convert = FALSE) %>%
        select(title, name, description) %>%
        drop_na(description) %>%
        group_by(title,name) %>%
        mutate(description_line = row_number()) %>%
        ungroup() %>%
        unnest_tokens(word, description) %>%
        inner_join(afinn) %>%
        group_by(title, index = description_line %/% 80) %>% 
        summarise(sentiment = sum(score))
      book_description_sentiments
      
      book_description_anticipation <- separate_rows(data_view, description, sep = "\\.", convert = FALSE) %>%
        mutate(description_length = str_length(description)) %>%
        filter(description_length > 0) %>%
        select(title, name, description) %>%
        drop_na(description) %>%
        group_by(title,name) %>%
        mutate(description_line = row_number()) %>%
        ungroup() %>%
        unnest_tokens(word, description) %>%
        inner_join(anticipation) %>%
        group_by(title, index = description_line %/% 80) %>% 
        summarise(anticipation = sum(score))
      book_description_anticipation
      
      book_description_scores <- merge(book_description_sentiments, book_description_anticipation, all = T)
      book_description_scores <- merge(book_description_scores, data_view, all = T)
      
      book_description_scores %>%
        mutate(title_lower = str_to_lower(title),
               title_formatted = str_replace_all(title_lower," ","+"),
               author_formatted = str_replace_all(str_to_lower(name)," ","+"),
               title = paste0('<a href="https://smile.amazon.com/s?k=',author_formatted,'+',title_formatted,'&i=stripbooks">',title,'</a>')) %>%
        rename(date = publishedDate,
               author = name) %>%
        select(author, title, date, anticipation, sentiment) %>%
        arrange(date)
    
    # data_view %>%
    #   select(name, title, publishedDate) %>%
    #   arrange(publishedDate)
  }
  else{
    paste0('No results found for ',input$author_query)
  }
}
# allow html hyperlinks
, sanitize.text.function = function(x) x)
```

### Word frequencies
  
```{r upcoming_books_cloud, fig.width=10, fig.height=10}
		
renderPlot({
  
  data_view <- getData()
  # for debug
  # data_view <- books
  
  if(length(data_view) > 0){
    data_view <- data_view
      
    descriptions <- unnest_tokens(data_view, word, description) %>%
      anti_join(stop_words) %>%
      count(name, title, word, sort = TRUE)
  
    if(length(descriptions) > 0){
      descriptions %>%
      with(wordcloud(word, n, max.words = 100, colors=wes_palette("Darjeeling1"), scale=c(3.0,0.5), rot.per=0.1))
    }
  }
})
```

Sentiments
============================================================================

Row
----------------------------------------------------------------------------

### Book description anticipation sentiments

```{r upcoming_book_descriptions_series_anticipation_sentiments, fig.height=13}

# Can we judge a book by its description?

renderPlot({
  
  data_view <- getData()
  # for debug
  # data_view <- books
  
  # anticipation score
  anticipation <- anticipationScores()
    
  if(length(data_view) > 0){
    
    book_descriptions <- separate_rows(data_view, description, sep = "\\.", convert = FALSE) %>%
      select(title, name, description) %>%
      drop_na(description) %>%
      group_by(title,name) %>%
      mutate(description_line = row_number()) %>%
      ungroup() %>%
      unnest_tokens(word, description) %>%
      inner_join(anticipation) %>%
      count(word, sentiment, sort = TRUE) 
    
    # check whether there are both positve and negative
    pos <- book_descriptions %>%
      filter(sentiment == 'positive')
    neg <- book_descriptions %>%
      filter(sentiment == 'negative')
    
    if((nrow(pos)>0) & (nrow(neg)>0))
    {
      book_descriptions <- book_descriptions %>%
        acast(word ~ sentiment, value.var = "n", fill = 0) %>%
        comparison.cloud(colors=c("#F98302","#5BBCD6"),
                       title.bg.colors=c("#F98302","#5BBCD6"),
                       scale=c(3.0,0.5), max.words = 100)
    }
    else
    {
      if(nrow(pos)>0){
        book_descriptions <- book_descriptions %>%
        with(wordcloud(word, n, max.words = 100, colors=c("#5BBCD6"),
                       scale=c(3.0,0.5), rot.per=0.1))
      }
      else
        {
          if(nrow(neg)>0){
          book_descriptions <- book_descriptions %>%
          with(wordcloud(word, n, max.words = 100, colors=c("#F98302"),
                         scale=c(3.0,0.5), rot.per=0.1))
          }
        }
    }
      
  }
  else{
    paste0('No results found for ',input$author_query)
  }
})
```

### Book description sentiments

```{r upcoming_book_descriptions_series_sentiments, fig.height=13}

# Can we judge a book by its description?

renderPlot({
  
  data_view <- getData()
  # for debug
  # data_view <- books
    
  if(length(data_view) > 0){
    
    book_descriptions <- separate_rows(data_view, description, sep = "\\.", convert = FALSE) %>%
      select(title, name, description) %>%
      drop_na(description) %>%
      group_by(title,name) %>%
      mutate(description_line = row_number()) %>%
      ungroup() %>%
      unnest_tokens(word, description) %>%
      inner_join(get_sentiments("bing")) %>%
      count(word, sentiment, sort = TRUE) %>%
      acast(word ~ sentiment, value.var = "n", fill = 0) %>%
      comparison.cloud(colors=c("#F98302","#5BBCD6"),
                       title.bg.colors=c("#F98302","#5BBCD6"),
                       scale=c(3.0,0.5), max.words = 100)
  }
  else{
    paste0('No results found for ',input$author_query)
  }
})
```

Raw Data
============================================================================

Row
----------------------------------------------------------------------------

### Upcoming book descriptions

```{r upcoming_book_descriptions}

# View all results, todo: maybe limit 1 description per publisher? or perhaps average scores

renderTable({
  
  data_view <- getData()
  # for debug
  # data_view <- books
  
  if(length(data_view) > 0){
    data_view <- data_view %>%
      drop_na(description) %>%
      select(name, title, publishedDate,description) %>%
      distinct(name, title, publishedDate,description) %>%
      arrange(publishedDate)
  }
  else{
    paste0('No results found for ',input$author_query)
  }
})
```

About
============================================================================

Row
----------------------------------------------------------------------------

### Motivations

As an avid reader of mysteries, I look forward to reading the latest books from my favorite authors. In order to stay informed of release dates, mybookQ queries [google books api](https://developers.google.com/books/) to [search](https://developers.google.com/books/docs/v1/using#PerformingSearch) by author name to return the newest results.

I also would like to have a good idea on the likelihood of enjoying a book because even my most favorite author will occasionally drift into murky and less cheerful mysteries or even worse, produce a stand-alone novel that includes none of my favorite fictional friends. By performing sentiment analysis as well as checking for specific key words, I will score an Anticipation Index for all not yet released books. It would also be helpful to have a forecast for the next year in books so that I know when I must seek out new authors recommended by the Seattle Times [Adam Woog](https://www.seattletimes.com/author/adam-woog/) and now digital [Seattle Mystery Books blog](https://mysteryish.wordpress.com/).

I believe that my favorite authors' stories are ones of great kindness to all, and if I do glean common language in descriptions and reviews, I would like to use that to help discover new authors rather than solely trust what other shoppers have bought like me, because I will make a bad decision and opt to neither finish nor review book choice mishaps.

- Thanks to Julia Silge and David Robinson's awesome book  [Text Mining with R: A Tidy Approach](https://www.tidytextmining.com)! Get it on [Amazon](https://smile.amazon.com/dp/B071K9RT9Z)!
- Thanks to Kyle Walker's [Neighborhood Diversity  in the American metropolis](https://walkerke.shinyapps.io/neighborhood_diversity/) Shiny application with the multiple pages layout. Amazing data exploration tool with just as awesome design! #GoFrogs!
- As always, thanks to [Karthik Ram](http://inundata.org/) for the pleasing [Wes Anderson color palettes](https://github.com/karthik/wesanderson)
- [Dean Attali](https://deanattali.com/)'s awesome [timevis](https://daattali.com/shiny/timevis-demo/) for the perfect timeline visualization

Happy Reading!