---
title: "Book Search Results Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=F)
```

# Initial Exploration and Anaylsis of Books by Author

## Exploratory Analysis
* Load up Search Results for Top 5 Authors
* Verify that results are for the expected authors
* Visualize books over time

```{r libraries}
library(ggraph)
library(igraph)
library(lubridate)
library(tidytext)
library(tidyverse)
library(widyr)
data("stop_words")
```

### The Data
Initially loaded `publishedData` as a Data, but received warning that 7 records contained only a year and not a full date, so will need to set those records to an arbitrary date like "1/1/year" for now.

```{r loadData, echo=F}
data_dir <- "/users/alicia.brown/repos/mybookq/data/"
books_file <- paste0(data_dir,"book_search_results.csv")
books_raw <- read_csv(books_file, col_types = "ccccccnncc", trim_ws = T)

authors_file <- paste0(data_dir,"books - authors.csv")
authors_raw <- read_csv(authors_file, col_types = "ccn")

summary(books_raw)
```

#### Clean Data
* Fix Dates that are years
* Filter by expected authors
* Set any missing (NAs) Average Ratings to Zero (for now) so all books are plotted
```{r fixDates}
fix_year_date <- function(published_date){
  if(str_length(published_date) == 4){
    return(paste0(published_date,"-01-01"))
  }
  else return(published_date)
}
# make_a_date <- fix_year_date("2017")
# already_a_date <- fix_year_date("2017-03-07")

books <- books_raw

books$publishedDate <- map(books$publishedDate, fix_year_date)
books$publishedDate <- ymd(books$publishedDate)

#set Name to lower case in prep for join to authors
books$name <- str_to_lower(books$authors)

# set NA Ratings to 0 so we don't exclude any books since the minimum ranking = 2.5
books$averageRating <- replace_na(books$averageRating, 0)

authors_raw$name <- paste(authors_raw$`first name`, authors_raw$`last name`)
books_authors <- inner_join(books, authors_raw,  by = c("name"))
summary(books_authors)
```
#### Visualize each Author by Average Rating over Time
* Are Louise Penny and Fred Vargas the highest ranked?
* Note that points with zero value are really missing Rating data
```{r vizBooksTimeline}
ggplot(books_authors) +
  geom_line(mapping = aes(x = publishedDate, y = averageRating,
                          color = name)) +
  geom_point(mapping = aes(x = publishedDate, y = averageRating,
                           color = name)) +
  facet_wrap(~ name, nrow = 5)
```

### Text Exploration

#### Explore book descriptions for Fred Vargas
```{r textVargas}
books_vargas <- books_authors %>%
  filter(name=="fred vargas") %>%
  arrange(publishedDate)

books_vargas[,c(4,2,9)]
```

#### Explore book descriptions for Louise Penny
```{r textPenny}
books_penny <- books_authors %>%
  filter(name=="louise penny") %>%
  arrange(publishedDate)

books_penny[,c(4,2,9)]
```

#### Create Document Matrix
```{r docPenny}
penny_description <- books_penny[,c(2,9)] %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words)

penny_description %>%
  count(word, sort = T)

penny_description %>%
  group_by(word) %>%
  count(word, sort = T)

penny_description %>%
  count(word, sort = T) %>%
  filter(n > 5) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) + 
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

```{r docVargas}

vargas_description <- books_vargas[,c(2,9)] %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words)

vargas_description %>%
  count(word, sort = T)

vargas_description %>%
  group_by(word) %>%
  count(word, sort = T)

# visualize top word frequency
vargas_description %>%
  count(word, sort = T) %>%
  filter(n > 4) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) + 
  coord_flip()
```

#### Term Frequency


```{r pennytfidf}
# term frequency - inverse document frequency
penny_description_tfidf <- penny_description %>%
  count(title, word, sort = T) %>%
  ungroup() %>%
  bind_tf_idf(word, title, n)

# highest tf-idf words in descriptions
penny_description_tfidf %>%
  arrange(-tf_idf)
```

```{r vargastfidf}
# term frequency - inverse document frequency
vargas_description_tfidf <- vargas_description %>%
  count(title, word, sort = T) %>%
  ungroup() %>%
  bind_tf_idf(word, title, n)

# highest tf-idf words in descriptions
vargas_description_tfidf %>%
  arrange(-tf_idf)
```

## Compare word frequency across both authors
```{r frequencyAuthors}
frequency <- bind_rows(mutate(penny_description, author = "Louise Penny"),
                       mutate(vargas_description, author = "Fred Vargas")) %>%
  mutate(word = str_extract(word, "[a-z']+")) %>%
  count(author, word) %>%
  group_by(author) %>%
  mutate(proportion = n / sum(n)) %>%
  select(-n) %>%
  spread(author, proportion) %>% 
  gather(author, proportion, `Louise Penny`:`Fred Vargas`)

cor.test(data = frequency[frequency$author == "Louise Penny",],
         ~ proportion + `Fred Vargas`)

frequency[frequency$author == "Louise Penny",]
frequency[frequency$author == "Fred Vargas",]
data = frequency[frequency$author == "Louise Penny",]

```